---
title: "Covid91 vaccine study Final2021"
author: "Isabella Gonzalez"
date: "Spring 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methods

The first step is isolating the subpopulations 
```{r}
maledata<-filter(FinalData,sex=="guy")
maledata$sex <- as.factor(maledata$sex)

```

# Males

## Barchart

```{r}
Males <- maledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n()) 

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
  
The bar chart shows that out of approximately 340 males who were infected with Covid-19, about 56% of those received the drug and 43% were given the placebo.

## Numerical Summary

```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)

```

This data shows that about 340 males were infected with Covid-19 out of approximately 10,400 males within the study. Furthermore 3.72% of the males who had received the drug got Covid-19 and 2.55% who received the placebo contracted Covid-19. Therefore this data is showing that males who received the vaccine were about 50% more likely to contract Covid suggesting that the drug is ineffective for males.

## Inferential  

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi square data shows that p-value of 0.0006181 is so small that it is highly unlikely the results were by chance.   
$H_0$ The null hypothesis is given that taking the vaccine makes no difference in the male population contracting Covid-19 or not.  
$H_A$ The alternative hypothesis is that the vaccine in fact does make a difference in whether or not a male contracts Covid-19. Therefore we reject the null hypothesis.

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odd ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 59.5% effectiveness.


# Females

```{r}
femaledata<-filter(FinalData,sex=="gal")
femaledata$sex <- as.factor(femaledata$sex)

```

## Barchart

```{r}
Females <- femaledata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n()) 

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

The bar chart shows that out of approximately 610 females who were infected with Covid-19, 32% of those received the drug and 66% were given the placebo.

## Numerical Summary

```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)

```

This data shows that about 610 females were infected with Covid-19 out of approximately 8,600 females within the study. Furthermore 4.44% of the females who had received the drug got Covid-19 and 8.67% who received the placebo contracted Covid-19. Therefore this data is showing that females who received the vaccine were about 50% less likely to contract Covid suggesting that the drug is effective for females.

## Inferential  

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi square data shows that p-value of 3.626e-16 is so small that it is highly unlikely the results were by chance.   
$H_0$ The null hypothesis is given that taking the vaccine makes no difference in the female population contracting Covid-19 or not.  
$H_A$ The alternative hypothesis is that the vaccine in fact does make a difference in whether or not a female contracts Covid-19. Therefore we do not reject the null hypothesis.

The odds ratio from the Fisher Exact test is 2.05:1, when taking the odd ratio and computing it into a percent of effectiveness by using the formula O/(1+0)*100 you get 67.16% effectiveness.

# LGBTQ


# Drug User

```{r}
Druggiedata<- filter(FinalData,DrugUser== "yes")
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)

```

## Barchart

```{r}
Druggies <- Druggiedata %>% group_by(infected, treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n()) 

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")


#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```
The bar chart shows that out of approximately 117 Druggies who were infected with Covid-19, 32% of those received the drug and 66% were given the placebo.

## Numerical Summary

```{r}
table3 <- xtabs(~infected + treatment, data=Druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)

```

This data shows that about 117 druggies were infected with Covid-19 out of approximately 2166 druggies within the study. Furthermore 3.63% of the druggies who had received the drug got Covid-19 and 6.46% who received the placebo contracted Covid-19. Therefore this data is showing that druggies who received the vaccine were about 50% less likely to contract Covid suggesting that the drug is effective for druggies.


# Discussion & Conclusion 

